---
title: "Utah Football - Simple Linear Regression"
output: 
  html_document:
    theme: cerulean
    code_folding: hide
---

```{r, message=FALSE, warning=FALSE}
# Load your libraries
library(car)
library(tidyverse)
library(mosaic)
library(pander)
library(DT)


utes <- read.csv("../../Data/utahfootball.csv", header=TRUE)
# or
#   library(readr)
#   someName <- read_csv("../Data/YourDataFileName.csv")

# Don't forget to run "Session -> Set Working Directory -> To Source file location"

```


<!-- Directions

Come up with a simple linear regression of your own creation. 

* Write the Background and clearly state your question. Then collect data and perform the linear regression analysis in order to answer your question. Use the proper order of first, ask the question, then second, figure out the answer.

* You should collect both the quantitative response variable and the quantitative explanatory variable yourself. This can be data you collect from 
    - the internet (like weather data, Facebook data, or sports data) 
    - your own experiment 
    - your own observational study
* However,     
    - The data should not be a single pre-existing dataset. It must be something that you piece together yourself.

* You need at least $n=3$ $(X,Y)$ pairs to perform the regression. However, your regression will only be useful if you can get around $n=10$ or more observations.

-->


## Background

Weather has an impact in almost everything we do, but does it have an impact on how many points are scored in a University of Utah football game? For this study, I have gather scores and average temperatures of all the Utah football games within the last five years. This will give us a total of 75 observations to compare. A sample of the data is shown in the table below.

```{r}
datatable(utes)
```

## Analysis

For this analysis, we will use the equation. 
$$
  \underbrace{Y_i}_\text{Final Score} = \beta_0 + \beta_1 \underbrace{X_i}_\text{Average Temperature} + \epsilon_i \quad \text{where} \ \epsilon_i \sim N(0, \sigma^2) 
$$
Where we'll be looking, in particular at the slope of the regression model, $\beta_1$. If the slope is zero, there is no relationship between the average actual score of the game and the reported score of game.

$$
  H_0: \beta_1 = 0
$$
$$
  H_a: \beta_1 \neq 0
$$
$$
  \alpha = 0.05
$$

The scatterplot below shows how Temperature plays a factor in Points scored in a football game which is only slight at a correlation of 0.1525(or the slope of the linear regression). The dots are also not close to the trend line showing that we have high variance. 

```{r}
plot(points_scored ~ avg_temperature, data=utes, pch=21, bg="blue", col="gray", 
     xlab="Average Temperature (Fahrenheit)", ylab="Points Scored", 
     main="Temperature vs Points in Football", las=1)
utes.lm <- lm(points_scored ~ avg_temperature, data=utes)
abline(utes.lm, lwd=3, col=rgb(.4,.4,.4,.2))
abline(v=seq(40,120,20), h=seq(40,120,20), lty=2, col=rgb(.6,.6,.6,.2))

```

A summary of the Regression analysis is shown below for your reference. 

```{r message=FALSE, warning=FALSE}
summary(utes.lm) %>% 
  pander()

```

## Appropriatenesss of the above Regression

Looking at whether or not this test was appropriate for this study, examine the three plots below. The red-line or linearity in the Residuals vs Fitted Values plot is flat enough to not raise any red flags but the vertical variation seems to be pretty inconsistent. The QQ-Plot is showing a lack of normality especially in the extreme values. The Residuals vs Order doesn't give us reason for concern as there seems to be no patterns or correlations in the dots.

```{r, fig.height=3}
par(mfrow=c(1,3))
plot(utes.lm, which=1:2)
plot(utes.lm$residuals, main="Residuals vs Order", xlab="",
     ylab="Residuals")
```

## Interpretation
The estimated slope being at 0.1525, we are awfully close to zero. Looking at the p-value confirms the visual given by the scatterplot with our p-value being p = 0.1453, which is greater than alpha. Therefore we do not have sufficient evidence to reject the Null Hypothesis. Showing us that the relationship between average temperature and points scored is not significant.

## Conclusion

Weather plays a role in many things; however, from the test performed, it does not play a role in whether or not the Utah Utes will score a certain amount of points in a football game. The test does not tell us whether or not weather plays a role in other factors of the game. This would be an interesting aspect to look at. 





























