---
title: "Skills Quiz: Different Types of Models"
output: 
  html_document:
    theme: cerulean
    code_folding: hide
    toc: true
    toc_float: true
---


## Instructions

Use this file to keep a record of your work as you complete the "Skills Quiz: Different Types of Models" assignment in Canvas.


```{r, message=FALSE, warning=FALSE}
library(mosaicData)
library(tidyverse)
```


----

<!-- Note: The {} after each Problem and Part header allows you to keep track of what work you have completed. Write something like {Done} once you complete each problem and your html file will then show you a nice summary of what you have "done" already. -->



## Problem 1 {}

Consider the scatterplot shown here of a single residence's monthly gas bill according to the month of the year. See `?Utilities` for more details on the data.

```{r}

plot(gasbill ~ month, data=Utilities, main="Single Residence in Minnisota", xlab="Month of the Year", ylab="Monthly Gas Bill (US Dollars)")

utilities.lm <- lm(gasbill ~ month + I(month^2), data = Utilities)

b <- utilities.lm$coefficients

curve(b[1] + b[2] * x + b[3] * x^2, col = "blue", add = T)


```



### Part (a)

Add the estimated regression function to the scatterplot above and state the function below.

<div class="YourAnswer">

$$
  \hat{Y}_i = `r b[1]` + `r b[2]`X_i + `r b[3]X_i^2
$$

</div>

### Part (b)

Diagnose the appropriateness of this regression model. How well does it fit the data? 

<span style="font-size:.8em;">Be sure to provide diagnostic plots and supporting arguments for your claims.</span>

<div class="YourAnswer">

```{r}
par(mfrow = c(1,3))
plot(utilities.lm, which = 1:2)
plot(utilities.lm$residuals)
```

</div>


### Part (c)

What range of possible gas bill amounts do you predict for the September bill? How confident are you in your prediction?

<div class="YourAnswer">

```{r}
predict(utilities.lm, data.frame(month = 9), interval = "prediction")
```

Type your answer here...

</div>


### Part (d)

Compute the mean of just September's gas bills and draw a horizontal reference line for this mean on your scatterplot. How does this mean compare to your prediction in Part (c)?

<div class="YourAnswer">

```{r}
sep_utilities <- Utilities %>% 
  filter(month == 9)

a <- sep_utilities %>% 
  summarize(mean = mean(gasbill))

38.84658 - a

```

Type your answer here...

</div>




### Part (e)

State the values of $R^2$, $MSE$, and the residual standard error. What do each of these values tell us about the regression model?

<div class="YourAnswer">

```{r}
summary(utilities.lm)
```

Type your answer here...

</div>

<br />



## Problem 2

View the `mtcars` dataset and corresponding help file `?mtcars`.

Perform a regression that predicts the miles per gallon `mpg` of the vehicle based on the quarter mile time `qsec` and transmission type `am` of the vehicle.


### Part (a)

Plot the data and your fitted regression model.

<div class="YourAnswer">

```{r}
par(mfrow = c(1,1))
plot(mpg ~ qsec, data = mtcars, col = as.factor(am))
mt.lm <- lm(mpg ~ qsec + am + qsec:am, data = mtcars)
b <- mt.lm$coefficients

x2=0
curve(b[1] + b[2]*x + b[3]*x2 + b[4]*x*x2, add = T, col = "black")
x2=1
curve(b[1] + b[2]*x + b[3]*x2 + b[4]*x*x2, add = T, col = "red")



```

</div>


### Part (b)

State the fitted regression model.

<div class="YourAnswer">

$$
 \underbrace{\hat{Y_i}}_\text{mean mpg} = `r b[1]` + `r b[2]`\underbrace{X_\text{1i}}_\text{qsec} + `r b[3]`X_{2i} + `r b[4]`X_\text{1i}X_\text{2i} + \epsilon_i
$$

</div>

### Part (c)

Use `summary(...)` to perform an appropriate t test to determine if the interaction term is needed in this regression model.

<div class="YourAnswer">

```{r}
summary(mt.lm)
```

Type your answer here...

</div>


### Part (d)

Diagnose the appropriateness of this regression model. How well does it fit the data? 

<span style="font-size:.8em;">Be sure to provide diagnostic plots and supporting arguments for your claims.</span>

<div class="YourAnswer">

```{r}
par(mfrow = c(1,3))
plot(mt.lm, which = 1:2)
plot(mt.lm$fitted.values)
```

Type your answer here...

</div>


### Part (e)

State the $R^2$ and residual standard error values for this model. What do these values show?

<div class="YourAnswer">

```{r}
summary(mt.lm)
```

Type your answer here...

</div>

<br />



## Problem 3

View the `mtcars` dataset and corresponding help file `?mtcars`.

Perform a regression that predicts the quarter mile time `qsec` of the vehicle based on the displacement of the engine `disp` and transmission type `am` of the vehicle according to the model:

$$
  \underbrace{Y_i}_\text{qsec} = \beta_0 + \beta_1 \underbrace{X_{1i}}_\text{disp} + \beta_2 X_{1i}^2 + \beta_3 \underbrace{X_{2i}}_\text{am == 1} + \beta_4 X_{1i}X_{2i} + \beta_5 X_{1i}^2X_{2i} + \epsilon_i
$$

where $\epsilon_i \sim N(0, \sigma^2)$.

### Part (a)

Plot the data and your fitted regression model.

<div class="YourAnswer">

```{r}
par(mfrow = c(1,1))
plot(qsec ~ disp, data = mtcars, col = as.factor(am))
mtq.lm <- lm(qsec ~ disp + I(disp^2) + am + disp:am + I(disp^2):am, data = mtcars)
b <- mtq.lm$coefficients

mtcars %>% 
  ggplot(aes(y = qsec, x = disp)) +
  geom_point(aes(color = as.factor(am))) +
  labs(x = "Displacement (disp)",
       y = "Quarter Mile Time (qsec)",
       title = "Quadratic Model using Mtcars Data set",
       color = "Transmission") +
  stat_function(fun = function(x) b[1] + (b[2]*x) + (b[3]*x^2), col = "red") +
  stat_function(fun = function(x) (b[1] + b[4]) + ((b[2]+b[5])*x) + ((b[3]+b[6])*x^2), col = "blue") +
  theme(plot.title = element_text(hjust = .5)) +
  scale_color_manual("Transmission", values = c("red","blue"), labels = c(
    "Automatic", "Manuel"))
  
  
```

Type your answer here...

</div>

### Part (b)

State the fitted regression model.

<div class="YourAnswer">

$$
  \hat{Y}_i = ...
$$

</div>


### Part (c)

Use `summary(...)` to perform appropriate t tests to determine which interaction terms are needed in this regression model.

<div class="YourAnswer">

```{r}
summary(mtq.lm)
```

</div>

### Part (d)

Diagnose the appropriateness of this regression model. How well does it fit the data? 

<span style="font-size:.8em;">Be sure to provide diagnostic plots and supporting arguments for your claims.</span>


<div class="YourAnswer">

```{r}
par(mfrow = c(1,3))
plot(mtq.lm, which = 1:2)
plot(mtq.lm$fitted.values)
```

Type your answer here...

</div>

### Part (e) 

Drop the "least significant" term from the model. This is a silly statement really, but is an accepted practice when searching for a "best" model in regression analysis. What has changed?

<div class="YourAnswer">

```{r}
plot(qsec ~ disp, data = mtcars, col = as.factor(am))
mtqr.lm <- lm(qsec ~ disp + I(disp^2) + am + I(disp^2):am, data = mtcars)
b <- mtqr.lm$coefficients
summary(mtqr.lm)
```

Type your answer here...

</div>


```{r}

n <- 100

for (i in n){
  Xi <- rep(seq(-2, 3, length.out=n/2), each=2)
  X2 <- sample(c(0,1),n, replace = T)
  Yi <- -2 + (3*Xi) + (4*Xi^2) + ((2-(-2))*X2)+ ((5-3)*Xi*X2) + ((-3-4)*Xi^2*X2) + rnorm(n, 0, 2)
}

sim_model.lm <- lm(Yi ~ Xi + I(Xi^2) + X2 + Xi:X2 + I(Xi^2):X2)
b <- sim_model.lm$coefficients

sim_model <- data.frame(x = Xi, y = Yi, z = X2)

sim_model %>% 
  ggplot(aes(x = x, y = y)) +
  geom_point(aes(color = as.factor(z))) +
  labs(x = "X values",
       y = "Simulated Y values",
       title = "Simulated Quadratic Model challenge") +
  stat_function(fun = function(x) b[1] + (b[2]*x) + (b[3]*x^2), col = "red") +
  stat_function(fun = function(x) (b[1] + b[4]) + ((b[2]+b[5])*x) + ((b[3]+b[6])*x^2), col = "blue") +
  stat_function(fun = function(x) -2 + (3*x) + (4*x^2), col = "red") +
  stat_function(fun = function(x) (2) + ((5)*x) + ((-3)*x^2), col = "blue") +
  theme(plot.title = element_text(hjust = .5)) +
  scale_color_manual("", values = c("red","blue"), labels = c(
      "y=-2+3x+4x^2", "y=2+5x-3x^2"))


```




```{r}
loblolly3.lm <- lm(height ~ age + I(age^2) + I(age^3) + I(age^4), data = Loblolly)
loblolly2.lm <- lm(height ~ age + I(age^2) + I(age^3), data = Loblolly)
loblolly1.lm <- lm(height ~ age + I(age^2), data = Loblolly)
loblolly4.lm <- lm(height ~ age + I(age^2) + I(age^3) + I(age^4) + I(age^5), data = Loblolly)

par(mfrow = c(1,4))
plot(loblolly3.lm, which = 1)
plot(loblolly2.lm, which = 1)
plot(loblolly1.lm, which = 1)
plot(loblolly4.lm, which = 1)



```









<style>

.YourAnswer {
  color: #317eac;
  padding: 10px;
  border-style: solid;
  border-width: 2px;
  border-color: skyblue4;
  border-radius: 5px;
}

</style>

 
 