---
title: "R Cheat Sheets & Notes"
output: 
  html_document: 
    theme: cerulean
    highlight: tango
    css: styles.css
---

## Cheat Sheets

* [R Colors](http://www.stat.columbia.edu/~tzheng/files/Rcolor.pdf)

* [R Base Graphics Cheat Sheet](http://www.joyce-robbins.com/wp-content/uploads/2016/04/BaseGraphicsCheatsheet.pdf)

* [R Base Commands Cheat Sheet](https://www.rstudio.com/wp-content/uploads/2016/10/r-cheat-sheet-3.pdf)

* [R Markdown Cheat Sheet](https://www.rstudio.com/wp-content/uploads/2015/02/rmarkdown-cheatsheet.pdf)

* [Keyboard Shortcuts](https://support.rstudio.com/hc/en-us/articles/200711853-Keyboard-Shortcuts)

## Notes

Add your own notes here...

```{r}
# Add R commands you want to remember here...

```

```{r}
# Or create other R chunks to put them in like this...

```

```{r, eval=FALSE}
# Using ```{r, eval=FALSE} turns off the chunk, but still shows it.
# Useful when you want to remember code, but not run it in this file.

#Step 1
mtcars1 <- filter(mtcars, cyl %in% c(4,8))
myTest <-  t.test(wt ~ cyl, data = mtcars1, mu = 0)
observedTestStat <- myTest$statistic
#Step 2
N <- 2000
permutedTestStats <- rep(NA, N)
for (i in 1:N){
  permutedData <- sample(mtcars1$wt)
  permutedTest <- t.test(permutedData ~ cyl, data = mtcars1, mu = 0)
  permutedTestStats[i] <- permutedTest$statistic
}
#Step 3
hist(permutedTestStats, col = "skyblue", xlim = c(-12,12))
abline(v=observedTestStat, col = 'red', lwd = 3)

sum(permutedTestStats > observedTestStat)/2000 * 2

```


```{r, eval=FALSE}
# Using ```{r, eval=FALSE} turns off the chunk, but still shows it.
# Useful when you want to remember code, but not run it in this file.

#Step 1
myTest <-  aov(correct ~ condition, data = Friendly)
observedTestStat <- summary(myTest)[[1]]$`F value`[1]
#Step 2
N <- 2000
permutedTestStats <- rep(NA, N)
for (i in 1:N){
  permutedData <- sample(Friendly$correct)
  permutedTest <- aov(permutedData ~ condition, data = Friendly)
  permutedTestStats[i] <- summary(permutedTest)[[1]]$`F value`[1]
}
#Step 3
hist(permutedTestStats, col = "skyblue", xlim = c(-1,20))
abline(v=observedTestStat, col = 'red', lwd = 3)

sum(permutedTestStats >= observedTestStat)/2000 
sum(permutedTestStats <= observedTestStat)/2000 

```


```{r, eval=FALSE}
# Using ```{r, eval=FALSE} turns off the chunk, but still shows it.
# Useful when you want to remember code, but not run it in this file.

#Step 1
myTest <-  glm(sat > 1000 ~ expend, data = SAT, family = binomial)
observedTestStat <- summary(myTest)[[12]][2,3]
#Step 2
N <- 2000
permutedTestStats <- rep(NA, N)
for (i in 1:N){
  permutedData <- sample(mtcars1$sat)
  permutedTest <- glm(sat > 1000 ~ expend, data = SAT, family = binomial)
  permutedTestStats[i] <- summary(myTest)[[12]][2,3]
}
#Step 3
hist(permutedTestStats, col = "skyblue", xlim = c(-5,0))
abline(v=observedTestStat, col = 'red', lwd = 3)

sum(permutedTestStats > observedTestStat)/2000 * 2

```



```{r, eval=FALSE}
# Using ```{r, eval=FALSE} turns off the chunk, but still shows it.
# Useful when you want to remember code, but not run it in this file.

set.seed(1140411)

sample1 <- rnorm(30, 69, 2.5)

sample2 <- rnorm(30, 69, 2.5)

theData <- data.frame(values = c(sample1,sample2), group = rep(c(1,2), each=30))

View(theData)

boxplot(values ~ group, data = theData)

#Step 1
myTest <-  t.test(values ~ group, data = theData, mu = 0)
observedTestStat <- myTest$statistic
#Step 2
N <- 2000
permutedTestStats <- rep(NA, N)
for (i in 1:N){
  permutedData <- sample(theData$group)
  permutedTest <- t.test(values ~ permutedData, data = theData, mu = 0)
  permutedTestStats[i] <- permutedTest$statistic
}
#Step 3
hist(permutedTestStats, col = "skyblue", xlim = c(-12,12))
abline(v=observedTestStat, col = 'red', lwd = 3)

sum(permutedTestStats >= observedTestStat)/2000
sum(permutedTestStats <= observedTestStat)/2000

```


```{r, EVAL = FALSE}
set.seed(121)
sample1 <- rnorm(30, 185, 8)
sample2 <- sample1 - rnorm(30, 0, 3.5)
theData <- data.frame(values = c(sample1,sample2), group = rep(c(1,2), each=30), id = rep(c(1:30),times=2))
View(theData)
with(theData, hist(values[group==1] - values[group==2]))

#Step 1
myTest <-  t.test(values ~ group, data = theData, paired = TRUE, mu = 0)
observedTestStat <- myTest$statistic
#Step 2
N <- 2000
permutedTestStats <- rep(NA, N)
for (i in 1:N){
  permutedData <- sample(x=c(1,-1),size =30, replace = TRUE)
  permutedTest <- with(theData, t.test(permutedData*(values[group==1] - values[group == 2]),  mu = 0))
  permutedTestStats[i] <- permutedTest$statistic
}
#Step 3
hist(permutedTestStats, col = "skyblue", xlim = c(-12,12))
abline(v=observedTestStat, col = 'red', lwd = 3)

sum(permutedTestStats >= observedTestStat)/2000
sum(permutedTestStats <= observedTestStat)/2000

```





